\MakeUppercase{Notes on Layered $H$-Partitions of $k$-Planar Graphs} \thanks{This work was partly funded by NSERC and MRI.}

### 2019-05-02 warmup.tex
Tiny change
Fixing....

  We make use of the fact that, if $G$ is a $k$-plane graph and $G^+$ is the plane graph obtained by adding a dummy vertex at each crossing in $G$, then $\dist_{G^+}(x,y)\le k\dist_G(x,y)$ for any $x,y\in V(G)$.

\begin{lem}\lemlabel{k-induction} The setup:
  \begin{compactenum}
    \item Let $G_0$ be a connected $k$-plane graph.
    \item Let $G=(V(G_0), E(G))$ be some subgraph of $G_0$ (so $E(G)\subseteq E(G_0)$).
    \item Let $G^+$ be the triangulation obtained by adding a vertex at each crossing point in $G_0$ and then adding additional edges to obtain a triangulation.
    \item Let $T$ be a BFS spanning tree of $G^+$ rooted at some vertex $r$.
    \item For each integer $i\ge 0$, let $V^+_i=\{v\in V(G^+):\dist_{G^+}(r,v)=i\}$, and let $V_i=V(G)\cap V^+_i$.
    \item Let $C$ be a cycle in $G^+$ with $r$ in the exterior of $C$ and such that
    \begin{compactenum} 
      \item No edge of $C$ is crossed by any edge of $G$; and
      \item $V(C)$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
      \begin{compactenum}
        \item $C[P_i]$ is a path; and
        \item $|P_i\cap V_j| \le 24k+4$ for all integers $j\ge 0$.
      \end{compactenum}
    \end{compactenum}
    \item Let $N$ and $N^+$ be the subgraphs of $G$ and $G^+$ consisting only of those edges and vertices contained in $C$ or the interior of $C$.
  \end{compactenum}
  Then $N$ has an $H$-partition $(B_x : x\in V(H))$ such that
  \begin{compactenum}
    \item for every integer $j\ge 0$, and all $x\in V(H)$, $|B_x\cap V_j|\le 24k+4$; 
    \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=B_{x_i}$; and
    \item $H$ has a tree decomposition whose largest bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
  \end{compactenum}
\end{lem}

@@

  Define a vertex set $X\subset V(G)$ as follows:  The set $X$ contains $V(Y^+)\cap V(G)\setminus V(C)$.  Next, for each vertex $v^+\in V(Y^+)\setminus V(G)$, we consider the two edges $vw,xy\in E(G_0)$ whose intersection is $v^+$.  Next we set
  \[   X_{vw} = \begin{cases} \{v,w\} & \text{if $vw\in E(G)$} \\
           \emptyset & \text{if $vw\not\in E(G)$,}
         \end{cases}
  \]
  \[   X_{xy} = \begin{cases} \{x,y\} & \text{if $xy\in E(G)$} \\
           \emptyset & \text{if $xy\not\in E(G)$} 
         \end{cases}
  \]
  and add $(X_{vw}\cup X_{xy})\setminus V(C)$ to $X$.

>>>>

Let $G$ be a $k$-planar graph.  We will use of the \emph{plane augmentation} $G^+$ of $G$ obtained by adding a \emph{dummy vertex} at each crossing in $G$, then adding a \emph{star vertex} $v_f$ in each face $f$ of the resulting graph and making $v_f$ adjacent to each vertex of $f$.

We will make use of the fact that $\dist_{G^+}(x,y)\le k\cdot\dist_G(x,y)$ for any $x,y\in V(G)$.  This follows immediately from the fact that any edge $xy\in E(G)$ has a corresponding path $x-y$ in $G^+$ of length at most $k$.  We will also make use of the fact that every dummy vertex of $G^+$ has degree 8.

\begin{lem}\lemlabel{k-induction} The setup:
  \begin{compactenum}
    \item Let $G$ be $k$-plane graph.
    \item Let $G^+$ be the plane augmentation of $G$.
    \item Let $T$ be a BFS spanning tree of $G^+$ rooted at some vertex $r$.
    \item For each integer $i\ge 0$, let $V^+_i=\{v\in V(G^+):\dist_{G^+}(r,v)=i\}$, and let $V_i=V(G)\cap V^+_i$.
    \item Let $C$ be a cycle in $G^+$ with $r$ in the exterior of $C$ and such that
    \begin{compactenum} 
      \item No edge of $C$ is crossed by any edge of $G$; and
      \item $V(C)$ can be partitioned into $P_1,\ldots,P_k$, $k\le 3$ such that for each $i\in\{1,\ldots,k\}$,
      \begin{compactenum}
        \item $C[P_i]$ is a path; and
        \item $|P_i\cap V_j| \le 24k+4$ for all integers $j\ge 0$.
      \end{compactenum}
    \end{compactenum}
    \item Let $N$ and $N^+$ be the subgraphs of $G$ and $G^+$ consisting only of those edges and vertices contained in $C$ or the interior of $C$.
  \end{compactenum}
  Then $N$ has an $H$-partition $(B_x : x\in V(H))$ such that
  \begin{compactenum}
    \item for every integer $j\ge 0$, and all $x\in V(H)$, $|B_x\cap V_j|\le 24k+4$; 
    \item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=B_{x_i}$; and
    \item $H$ has a tree decomposition whose largest bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
  \end{compactenum}
\end{lem}

@@

  Define a vertex set $X\subset V(G)$ as follows:  The set $X$ contains $V(Y^+)\cap V(G)\setminus V(C)$.  Next, for each dummy vertex $v^+\in V(Y^+)\setminus V(G)$, we consider the two edges $vw,xy\in E(G_0)$ whose intersection is $v^+$.  These edges correspond to two paths $v\rightsquigarrow w$ and $x\rightsquigarrow y$.  We collect the set $A_{x^+}$ of all vertices of $G^+$ that are on any face that contains any dummy vertex in $v\rightsquigarrow w$ or $x\rightsquigarrow y$.  We add the set $A_{x^+}\cap V(N^+)\setminus V(C)$ to $X$.

  Note that $A_{x^+}$ contains $\{v,w,x,y\}$ and that, for any vertex $z\in A_{x^+}$, $\dist_{G^+}(z,x^+) \le k$.  Careful counting also reveals that the size of $A_{x^+}$ is at most .
